Adaptive Video Streaming with Optimized Bitstream Extraction and PID-based Quality Control

### 2015-11-25 tmm.tex
Change encoding to UTF-8
Add pseudo code for bitstream extraction algorithm

\begin{figure}[h]
\centering
\includegraphics[width = 0.9\linewidth]{Bitstream-Extraction.jpg}
\caption{Bitstream extraction of SVC \label{fig:Bitstream-Extraction}}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width = 0.9\linewidth]{GOPStructure.jpg}
\caption{Group of pictures (GOP) in SVC \label{fig:GOP_Structure}}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width = 0.9\linewidth]{ModelVerification.png}
\caption{Comparison of actual and estimated distortion for sequence {\em Foreman} \label{fig:model_verification}}
\end{figure}

@@

We simulate a process of discarding all the enhancement packets from a stream, using the above greedy strategy. During this process, the discarding order of a packet is regarded as its priority and recorded. Then the packets' priority values can be stored in the bitstream (either in the priority id field of the NALU header or in separate SEI messages) for real extraction in the future.

The detailed discarding and priority assignment process is described below:
\begin{description}
  	\item[1)] Initially, the sequence error vector, $e_{seq}$, is set to $e_{full}$.
  	\item[2)] Within the packets that can be discarded, find the packet $m$ having the least RD impact on the sequence, \textit{i.e. }
	\begin{equation}
	\label{eq:R-D_impact_m}
	\Phi_m = \min_{i \in I_{top}} \Phi_i \: ,
	\end{equation}
	where $I_{top}$ stands for the top layer packets, those currently able to be discarded. The RD impact of a packet is calculated based on (\ref{eq:rd_impact}), as follows: 
	\begin{equation}
	\label{eq:R-D_impact_i}
	\Phi_i = \dfrac{\partial D}{\partial R} = \dfrac{PSNR(e_{seq}) - PSNR(e_{seq} + e_i)}{SIZE(i)} \: ,
	\end{equation}
	where $PSNR(*)$ means to obtain PSNR from the error vector, and $SIZE(i)$ is the size of packet $i$.
  	\item[3)]The packet, $m$, is then removed from the bitstream, and its error vector is added to the sequence's error vector, \textit{i.e.}
  	\begin{equation}
  	\label{eq:error_update}
  	e_{seq} = e_{seq} + e_m \: .
  	\end{equation}
  	\item[4)]The process from 2) to 3) is repeated until all enhancement packets are discarded, and every packet is assigned a priority value according to the removing order.
\end{description}

For long video sequences, an optimization window is usually needed, and the priority assignment and bitstream extraction is conducted independently in each window. The computational complexity of this priority assignment algorithm is $O(K^2)$, with $K$ being the number of frames involved in the priority assignment process, \textit{i.e.}, the size of the optimization window. Note that this computation complexity is negligible compared with the decoding process needed in the error vector acquisition.

@@

\bibitem{Bouten14}
N. Bouten, S. Latré, J. Famaey, W. V. Leekwijck, and F. D. Turck, ``In-Network Quality Optimization for Adaptive Video Streaming Services'', in {\em IEEE Transactions on Multimedia}, vol. 16, no. 8, pp. 2281-2293, 2014.

>>>>

\begin{figure}[h]
\centering
\includegraphics[width = 1.0\linewidth]{Bitstream-Extraction.jpg}
\caption{Bitstream extraction of SVC \label{fig:Bitstream-Extraction}}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width = 1.0\linewidth]{GOPStructure.jpg}
\caption{Group of pictures (GOP) in SVC \label{fig:GOP_Structure}}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width = 1.0\linewidth]{ModelVerification.png}
\caption{Comparison of actual and estimated distortion for sequence {\em Foreman} \label{fig:model_verification}}
\end{figure}

@@

We simulate a process of discarding all the enhancement packets from a stream, using the above greedy strategy. During this process, the discarding order of a packet is regarded as its priority and recorded. Then the packets' priority values can be stored in the bitstream (either in the priority identifier field of the NALU header or in supplementary enhancement information messages) for real extraction in the future.

The detailed discarding and priority assignment process is described below:
\begin{description}
\item[1)] Initially, the sequence error vector, $e_{seq}$, is set to $e_{full}$.
\item[2)] Within the packets that can be discarded, find the packet $m$ having the least RD impact on the sequence, \textit{i.e. }
\begin{equation}
\label{eq:R-D_impact_m}
\Phi_m = \min_{i \in I_{top}} \Phi_i \: ,
\end{equation}
where $I_{top}$ stands for the top-layer packets, those currently able to be discarded. The RD impact of a packet is calculated based on (\ref{eq:rd_impact}), as follows: 
\begin{equation}
\label{eq:R-D_impact_i}
\Phi_i = \dfrac{\partial D}{\partial R} = \dfrac{PSNR(e_{seq}) - PSNR(e_{seq} + e_i)}{SIZE(i)} \: ,
\end{equation}
where $PSNR(*)$ means to obtain PSNR from the error vector, and $SIZE(i)$ is the size of packet $i$.
\item[3)]The packet, $m$, is then removed from the bitstream, and its error vector is added to the sequence's error vector, \textit{i.e.}
\begin{equation}
\label{eq:error_update}
e_{seq} = e_{seq} + e_m \: .
\end{equation}
\item[4)]The process from 2) to 3) is repeated until all enhancement packets are discarded, and every packet is assigned a priority value according to the removing order.
\end{description}

For long video sequences, an optimization window containing a certain number of frames is usually needed, and the priority assignment and bitstream extraction is conducted independently in each window. Suppose the size of the optimization window is $K$, i.e., there are $K$ frames involved in the above priority assignment process, and each frame contains $L_Q$ quality layers (one base layer and $L_Q-1$ enhancement layers). Then the number of top layer packets in step 2) of the process is $K$, and the total number of packets that can be discarded is $K \cdot (L_Q-1)$. Given that, the priority assignment process in each optimization window can be more specifically described in Algorithm \ref{algo:greedy}. It is easy to see that the computational complexity of this priority assignment algorithm is $O(K^2)$. Note that this computational complexity is negligible compared with the decoding process needed in the error vector acquisition.

\begin{algorithm}
\caption{Greedy-like bitstream extraction algorithm}
\label{algo:greedy}
\begin{algorithmic}
    \STATE Set $e_{seq}$ to $e_{full}$, and set $p$ to 0.
    \WHILE{$p < K \cdot (L_Q - 1)$}
	    \FOR{$K$ top layer packets}
		    \STATE Find the packet $m$ which has least RD impact
		\ENDFOR
		\STATE Set the priority value of packet $m$ to $p$
		\STATE Remove $m$ from the sequence
		\STATE Set $e_{seq}$ to $e_{seq} + e_m$, and set $p$ to $p+1$
    \ENDWHILE
\end{algorithmic}
\end{algorithm}

A larger optimization window means each time we have more packets to compare and determine their priorities. Therefore, choosing a larger $K$ will result in more optimized extraction. However, doing this will also increase the computational complexity as well as the memory used to store the error vectors. Besides, if the 6-bit priority identifier field of the NALU header is used to store the priority values, only a range of 0$\sim$63 is supported, which also limits the value of $K$.

@@

\bibitem{Bouten14}
N. Bouten, S. Latre, J. Famaey, W. V. Leekwijck, and F. D. Turck, ``In-Network Quality Optimization for Adaptive Video Streaming Services'', in {\em IEEE Transactions on Multimedia}, vol. 16, no. 8, pp. 2281-2293, 2014.
